---
title: "Spatial Statistics Report"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
params:
  csd:
    label: "Consolidated data"
    value: NULL
  csd_path:
    label: "Path to Consolidated_data.txt"
    value: C:/Research/phenoptrExamplesData/200601 GC/Consolidated_data.txt
    input: file
  export_path: 
    label: "Path to directory with component data"
    value: C:/Research/phenoptrExamplesData/200601 GC/export
  from_phenotype:
    label: "'From' phenotype"
    value: "CD8+"
  to_phenotype:
    label: "'To' phenotype"
    value: "CK+"
  tissue_categories:
    label: "Tissue categories"
    value: NULL
vignette: >
  %\VignetteIndexEntry{Spatial Statistics Report}
---

<style type="text/css">
body {
    max-width: 800px;
}

table {
    width: auto !important;
    max-width: 100%;
    margin-bottom: 20px;
    margin-left: 0px;
}

img {
  border: none;
}
</style>

```{r setup, echo=FALSE,include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo=FALSE,fig.width=9, fig.height=6, 
                      comment=NA, warning=FALSE, message=FALSE)

suppressPackageStartupMessages(library(tidyverse))

# Make local vars for params, for easier debugging
csd = params$csd
csd_path = params$csd_path
export_path = params$export_path
from_phenotype = params$from_phenotype
to_phenotype = params$to_phenotype
tissue_categories = params$tissue_categories

phenos = phenoptr::parse_phenotypes(from_phenotype, to_phenotype)
from_name = names(phenos)[[1]]
to_name = names(phenos)[[2]]
```

```{r data}
if (is.null(csd))
  csd = phenoptr::read_cell_seg_data(csd_path)
field_col = phenoptr::field_column(csd)
fields = sort(unique(csd[[field_col]]))
plot_rows = ceiling(length(fields)/2)

category_names = ifelse(is.null(tissue_categories), "All",
  paste(tissue_categories, collapse=', '))

# Summary of fields and cell counts
cell_counts = csd %>% 
  group_by(!!field_col) %>% 
  nest(data=-!!field_col) %>% 
  mutate(summary=map(data, 
           function(d) map_dfc(phenos, 
                               ~sum(phenoptr::select_rows(d, .x))))) %>% 
  select(-data) %>% 
  unnest('summary')
```

### Summary

Cell data from `` `r csd_path` ``.  
_From_ phenotype: `r from_name`  
_To_ phenotype: `r to_name`  
_Tissue categories_: `r category_names`  

#### Summary of fields and cell counts

`r knitr::kable(cell_counts)`

```{r funcs}
# Make a ppp for a single field with the given phenotypes and tissue categories
make_ppp2 = function(field) {
  pp1 = phenoptr::make_ppp(csd, export_path, from_phenotype, field,
                           tissue_categories=tissue_categories)
  pp2 = phenoptr::make_ppp(csd, export_path, to_phenotype, field,
                           tissue_categories=tissue_categories)
  spatstat::superimpose(pp1, pp2)
}
```

### Nearest neighbor function G from `r from_name` to `r to_name`

The G function shows the distribution of nearest neighbors. 
It is the cumulative distribution function of the distance from
a typical cell to its nearest neighbor.

```{r Gcross, fig.height=plot_rows*4}
# Compute and plot Gcross for each field
par(mar=rep(4, 4), mfrow=c(plot_rows, 2))

purrr::walk(fields, function(field) {
  pp = make_ppp2(field)
  spatstat::plot.fv(spatstat::Gcross(pp, from_name, to_name),
                    main=field)
})
```

### L function from `r from_name` to `r to_name`

The L function is a normalized summary of the number of cells within
a radius of cells of another type.

These plots show L(r)/r vs r.

```{r Lcross, fig.height=plot_rows*4}
# Compute and plot Lcross for each field
par(mar=rep(4, 4), mfrow=c(plot_rows, 2))

purrr::walk(fields, function(field) {
  pp = make_ppp2(field)
  spatstat::plot.fv(spatstat::Lcross(pp, from_name, to_name),
                    ./r ~ r, main=field)
})
```

### Pair correlation from `r from_name` to `r to_name`

Pair-correlation summarizes the relative density of cells of one type 
at each distance from cells of a second type.

```{r pcfcross, fig.height=plot_rows*4}
# Compute and plot pcfcross for each field
par(mar=rep(4, 4), mfrow=c(plot_rows, 2))

purrr::walk(fields, function(field) {
  pp = make_ppp2(field)
  spatstat::plot.fv(spatstat::pcfcross(pp, from_name, to_name, divisor='d'),
                    main=field)
})
```

### References

A readable and comprehensive discussion of spatial statistics may
be found in Baddeley, Rubak and Turner, 
_Spatial Point Patterns: Methodology and Applications with R_, 
Chapman and Hall/CRC Press, 2016
([link](http://www.crcpress.com/books/details/9781482210200/)).

More information about the specific functions shown in this report is
available in the R help for `spatstat::Gcross()`, `spatstat::Lcross()`
and `spatstat::pcfcross()` and at the 
`spatstat` [website](https://spatstat.org/).

This report created by phenoptr `r packageVersion("phenoptr")`
and phenoptrReports `r packageVersion("phenoptrReports")` on `r Sys.Date()`.  
http://akoyabio.github.io/phenoptrReports
